### 2014-02-10 Beacon P3 Proposal/main.tex
update proposal

Real world applications of evolutionary optimization are often
hindered by the need to determine problem specific parameter
settings.  
Parameters can have a significant impact on search efficiency~\cite{goldberg:1991:gasize}
and finding good settings can be very time consuming~\cite{grefenstette:1986:optimalga}.
While some previous methods have reduced~\cite{Back:1992:selfadapt}
or removed~\cite{pelikan:2004:parameterlesshboa} the
need for parameter tuning, many do so by trading efficiency for
general applicability.  The Parameter-less Population Pyramid (P3)
is an evolutionary technique that requires \textbf{no parameters}
and is still broadly effective. P3 strikes a balance between
continuous integration of diversity and exploitative elitist
operators, allowing it to solve easy problems quickly and hard
problems eventually.  Work currently under review for publication
compared P3 with three optimally tuned state-of-the-art algorithms~\cite{thierens:2013:ltgahiff,doerr:2013:lambdalambda}.
Across seven problem classes P3 always found the
optimum at least a constant factor \emph{\textbf{faster}} than all comparisons.
More importantly, on the three most challenging problem classes,
P3 had a lower order of computational complexity as measured by
evaluations.  Based on over one trillion evaluations, these results
suggest P3 has wide applicability to a broad class of
problems.  \textbf{Therefore, the primary focus of this project is to integrate
P3 into X-TOOLSS in order to (1) allow non-experts easy access to P3,
(2) improve P3's ability to optimize real world problems and (3) expand
P3 to new evolutionary domains utilizing Genetic Programming.}

@@

A challenge for all theory based optimization techniques is to make the transition
to solving meaningful problems.  While the field has progressed significantly beyond
its origins, many practitioners use outdated evolutionary systems~\cite{bechmann:2013:hummies}
or rely on methods tailored to their specific problem~\cite{izzo:2013:hummies}.
The fact that these suboptimal optimizers can already outperform human designers
suggests that bridging the gap between academia and industry may yield spectacular results.

@@

Adam has graciously worked with us to perform these preliminary experiments, but more general
release is currently hindered by the expert centric design of P3's software.  Even now that Adam
knows how to use our software, it requires him to have two separate tool sets for optimization.
Therefore we feel it is imperative to integrate P3 into X-TOOLSS to make it more user friendly,
allowing more engineers to benefit from P3's quality.

Beyond accessibility, X-TOOLSS provides an invaluable potential to P3 for academic data collection.
The framework's optimization analysis tools have been refined through continuous multi-user use, and
will likely lead to further insights into P3's optimization.  Furthermore, as X-TOOLSS has a number
of built in evolutionary optimizers, it provides a perfect device for gathering comparative results
as P3 is applied to new and challenging problems.  Once integrated, non-computer scientists can utilize
X-TOOLSS to compare P3 directly with existing evolutionary techniques.  This can provide the foundation
for collaborative research between industry and academia, with industry gaining high quality solutions
to challenging problems and academia gathering data on how to improve optimization techniques.

@@

P3 is currently limited to evolving discrete value encoded linear genomes. 
Therefore we propose to leverage Grammatical Evolution~\cite{oneill:2001:grammatical}
to give P3 the ability to evolve entire programs.  Grammatical Evolution (GE) is a subset
of Genetic Programming which utilizes a non-trivial genotype to phenotype mapping, converting
binary strings into full fledged programs.  To apply GE, a domain expert first creates a
grammar which defines the structure of all possible useful programs.  When evaluating a
solution, bits are read from the individual to determine which grammar expansion rules to perform,
deterministically creating a program.  This is an ideal transition for P3 as from an
evolutionary standpoint GE applications will be no different than binary string applications as
only the evaluation step will have to change.

>>>>

Real world applications of evolutionary optimization are often
hindered by the need to determine problem specific parameter
settings.  
Parameters can have a significant impact on search efficiency~\cite{goldberg:1991:gasize}
and finding good settings can be very time consuming~\cite{grefenstette:1986:optimalga}.
While some previous methods have reduced~\cite{Back:1992:selfadapt}
or removed~\cite{pelikan:2004:parameterlesshboa} the
need for parameter tuning, many do so by trading efficiency for
general applicability.  The Parameter-less Population Pyramid (P3)
is an evolutionary technique that requires \textbf{no parameters}
and is still broadly effective. P3 strikes a balance between
continuous integration of diversity and exploitative elitist
operators, allowing it to solve easy problems quickly and hard
problems eventually.  Work currently under review for publication
compared P3 with three optimally tuned state-of-the-art algorithms~\cite{thierens:2013:ltgahiff,doerr:2013:lambdalambda}.
Across seven problem classes P3 always found the
optimum at least a constant factor \emph{\textbf{faster}} than the
comparison algorithms.
More importantly, on the three most challenging problem classes,
P3 had a lower order of computational complexity as measured by
evaluations.  Based on over one trillion evaluations, these results
suggest P3 has wide applicability to a broad class of
problems.  \textbf{Therefore, the primary focus of this project is to:
(1) integrate P3 into X-TOOLSS so it may be applied to
new, practical problems, (2) use feedback from P3 X-TOOLSS applications
to improve P3 and (3) expand
P3's representation to Genetic Programming and explore its
applicability to different kinds of problems.}

@@

A challenge for many optimization techniques is to make the transition
to solving meaningful problems.  While the field has progressed significantly beyond
its origins, many practitioners use outdated evolutionary systems~\cite{bechmann:2013:hummies}
or rely on methods tailored to their specific problem~\cite{izzo:2013:hummies}.
The fact that these suboptimal optimizers can already outperform human designers
suggests that bridging the gap between academia and industry may yield spectacular results.

@@

Adam has graciously worked with us to perform these preliminary
experiments, but P3's present form requires more work to setup a
problem than is likely tolerable by a typical domain user. 
%more general
%release is currently hindered by the expert centric design of P3's software.  Even now that Adam
%knows how to use our software, it requires him to have two separate tool sets for optimization.
Therefore we feel it is imperative to integrate P3 into X-TOOLSS to make it more user friendly,
allowing more engineers to benefit from P3's capabilities.

Beyond accessibility, X-TOOLSS provides an invaluable potential for data collection.
The framework's optimization analysis tools have been refined through continuous multi-user use, and
will likely lead to further insights into P3's performance and applicability.  Furthermore, as X-TOOLSS has a number
of built in evolutionary optimizers, it provides a perfect device for gathering comparative results
as P3 is applied to new problems.  Once integrated, domain experts can utilize
X-TOOLSS to compare P3 directly with existing evolutionary techniques.
This will provide a foundation
for collaborative research between industry and academia, with industry gaining high quality solutions
to challenging problems and academia gathering data on how to improve optimization techniques.

@@

P3 is currently limited to evolving discrete value encoded linear
genomes which also limits the range of problems the approach can address.
We therefore also propose to examine ways to expand P3 into the
Genetic Programming area. We plan to leverage Grammatical Evolution~\cite{oneill:2001:grammatical}
to give P3 the ability to evolve programs.  Grammatical Evolution (GE) is a subset
of Genetic Programming which utilizes a non-trivial genotype to phenotype mapping, converting
binary strings into full fledged programs.  To apply GE, a domain expert first creates a
grammar which defines the structure of all possible useful programs.  When evaluating a
solution, bits are read from the individual to determine which grammar expansion rules to perform,
deterministically creating a program.  This is an ideal transition for
P3 since GE applications will be appear as binary string
applications requiring 
only modificiation to the evaluation step.
